---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

Jumping Rivers R Package Repository
-----------------------------------

This is an R package repository for the [Jumping Rivers](www.jumpingrivers.com) training courses. While this repository is public, it's primarily purpose is internal to Jumping Rivers.

### Creating a new package

-   Create a new [repository](https://github.com/jr-packages) with the correct package name: `jrXxxx`
-   On travis:
    -   add the secure variable `GITHUB_PAT` (ask Colin). The `GITHUB_PAT` variable is needed to push packages to this drat repo.
    -   ensure [travis](https://travis-ci.org/profile/jr-packages) will build your package. You'll need to sync your account to pick up the package.
-   Clone the directory
-   Copy the contents of the `jrXxxx` directory to a new directory
    -   Replace XXX in the DESCRIPTION file and in the `R/jrXxx-package.R` file
    -   Update README.md
-   Push to GitHub
-   Add your package to the list below.
    -   Add package name to the correct vector in the README.Rmd file and knit the file
    -   (Please keep it in alphabetical order)


### Current Packages

```{r, echo = FALSE}
library(knitr)
library(tibble)
pkgs = c("jrAdvanced", "jrAnalytics", "jrAutomate", "jrBig", "jrEfficient", "jrGgplot2", "jrIntroBio", "jrIntroduction", "jrModelling", "jrNotes", "jrPredictive", "jrProgBio","jrProgramming", "jrPresentation", "jrShiny", "jrSql", "jrStan", "jrTidyverse", "jrTidyverse2","jrWhyR")
Packages = paste0("[", pkgs, "](https://github.com/jr-packages/", pkgs, ")")
Build_Status = paste0("[![Build Status](https://api.travis-ci.org/jr-packages/",
                      pkgs, ".png?branch=master)](https://travis-ci.org/jr-packages/)")
tib = tibble(Package = Packages, `Build Status` = Build_Status)
kable(tib)
```

### Other R Packages

```{r, echo = FALSE}
opkgs = c("efficientTutorial")
oPackages =  paste0("[", opkgs, "](https://github.com/jr-packages/", opkgs, ")")
oBuild_Status = paste0("[![Build Status](https://api.travis-ci.org/jr-packages/",
                      opkgs, ".png?branch=master)](https://travis-ci.org/jr-packages/)")
tib2 = tibble(Packages = oPackages, `Build Status` = oBuild_Status)
kable(tib2)
```

### To do 

Scrape package names automatically